---
title: Seasonal Adjustment Workshop Story
subtitle: for practicioners of seasonal adjustment
author: James Livsey
abstract: 
  This is an X-13 story. A seasonal adjustment
  discussions that can be turned into **PDFs** or into **interactive websites**.
output: x13story::x13story
---


```{r, echo = FALSE, message = FALSE, warning=FALSE}
# landscape aspect ratio for time series
knitr::opts_chunk$set(fig.width = 10, fig.height = 6)
# also load seasonal
library(x13story)  
```

## Let's get started using `x13story`

It is quite easy to communicate about your project in this format. This text
will be rendered exactly as seen. The beauty of this markdown language is that
it is compiled with R. This allows us to use R functions to manipulate our 
output!

While documenting our seasonal adjustment project, we can run X-13ARIMA-SEATS
from the same document...


```{r, echo = FALSE, fig.cap = "A first X-13 view."}
m <- seas(AirPassengers)

# the 'x13view' function initializes an interactive X-13 view, or adds a 
# graph to a PDF. 
x13view(m, "main")
```

### A default adjustment of the AirPassengers Series

Notice the seasonally adjusted sereis superimposed on the original series. 

### Data can be included in this file too

No longer will you need to send multiple files along with your project. The 
function `dput()` allows users to include data directly into this document.

```
dput(myseries)

``` 
```{r, echo = FALSE, fig.cap = "A second X-13 view."}
myseries <- structure(c(112, 118, 132, 129, 121, 135, 148, 148, 136, 119, 
55, 55, 55, 55, 55, 135, 125, 149, 170, 170, 158, 133, 114, 
140, 145, 150, 178, 163, 172, 178, 199, 199, 184, 162, 146, 166, 
171, 180, 193, 181, 183, 218, 230, 242, 209, 191, 172, 194, 196, 
196, 236, 235, 229, 88, 264, 272, 237, 211, 180, 201, 204, 188, 
235, 227, 234, 264, 302, 293, 259, 229, 203, 229, 242, 233, 267, 
269, 270, 315, 364, 347, 312, 274, 237, 278, 284, 277, 317, 313, 
318, 374, 413, 405, 355, 306, 271, 306, 315, 301, 356, 348, 355, 
422, 465, 467, 404, 347, 305, 336, 340, 318, 362, 348, 363, 435, 
491, 505, 404, 359, 310, 337, 360, 342, 406, 396, 420, 472, 548, 
559, 463, 407, 362, 405, 417, 391, 419, 461, 472, 535, 622, 606, 
508, 461, 390, 432), .Tsp = c(1949, 1960.91666666667, 12), class = "ts")

m <- seas(myseries, x11 = "")

# initializes a second X-13 view
x13view(m, 'x11.seasonal')
```

### An x11 adjustment of myseries

This is the seasonal component after adjusting myseries. 

### Latex 

You can also use \LaTeX-style math both for PDFs and and the interactive view:

$$X_t = T_t + S_t + I_t$$ 

Inlined math, $E = mc^2$, is possible as well.

You can include your own data into the code. Use `dput` to transform an existing series in your workspace into R code:




